---
title: '&#32;'
author: "&#32;"
date: "&#32;"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options("scipen" = 999)
```

<br>

# A Clever Title About Museum Grants

#### And a descriptive subtitle

------------------------------------------------------------------------

This project was completed during a course called "Design Lab: The Industry Challenge" as part of [MICA's Data Analytics and Visualization program](https://online.mica.edu/dav/). The goal of the project is to take students through the design process from start to finish, which not only includes working with data and creating visualizations, but also iterating on the design by incorporating feedback from peers and instructors. Students were required to use data from the Urban Institute, but the topic itself was open-ended.

I chose to explore federal grants made to museums and similar institutions in the US. Partly because the data runs the gamut in terms of numerical, categorical, geospatial, etc., but mostly because I am a museum nerd. I have always enjoyed exploring new museums in new places and used this project as an opportunity to explore them from a different perspective.

Here is a quick summary of what I learned:

-   Placeholder

-   Placeholder

-   Placeholder

<br>

*Note - I'll be using the term "museum" as a catch-all for museums, art galleries, botanical gardens, arboretums, nature centers, zoos, aquariums, and the like.*

------------------------------------------------------------------------

<br>

### Museum grants over the last 10 years

The Urban Institute led me to data from the Institute of Museum and Library Services (IMLS), an independent agency of the federal government on a mission to, "advance, support, and empower America's museums, libraries, and related organizations through grantmaking, research, and policy development." The IMLS maintains a list of museums and associated information, last updated in 2018. The IMLS also makes available data regarding the grants they award to museums through various programs.

Although the grant data goes as far back as 1996 I decided to bound the time frame to the last 10 years (2013 - 2022). Over that time period the IMLS **awarded \$101,674,483 through 628 grants** to museums in the US. The amount awarded each year see-sawed from 2013 to 2018 before **rising sharply in 2019**. It continues to increase through 2022, though at a slower rate.

I'll unpack this further in the next section.

<br>

```{r include=FALSE}

# load packages
library(tidyverse)
library(maps)
library(patchwork)
library(kableExtra)

# import grant data (v2)
data_grants_raw <- read_csv("awarded-grants-2023-04-01 v2.csv")

# select columns to work with
data_grants <- data_grants_raw %>%
  select(institution,
         city,
         state,
         year,
         funds,
         program,
         `log number`
         )

# import MDF File 1 (v2), format select columns
data_museums_raw <- read_csv("MuseumFile2018_File1_Nulls v2.csv", 
                             col_types = cols(DISCIPL = col_factor(levels = c("ART", "BOT", 
                                                                              "CMU", "HST", 
                                                                              "NAT", "SCI", 
                                                                              "ZAW")), 
                                              TAXPER15 = col_date(format = "%Y%m"),
                                              INCOMECD15 = col_factor(levels = c("0", "1", 
                                                                                 "2", "3", 
                                                                                 "4", "5", 
                                                                                 "6", "7",
                                                                                 "8", "9")),
                                              AAMREG = col_factor(levels = c("1", "2", "3",
                                                                             "4", "5", "6")),
                                              BEAREG = col_factor(levels = c("1", "2", "3",
                                                                             "4", "5", "6",
                                                                             "7", "8")),
                                              LOCALE4 = col_factor(levels = c("1", "2", "3", 
                                                                              "4"))
                                              )
                             )

# select columns to work with
data_museums <- data_museums_raw %>%
  select(COMMONNAME,
         DISCIPL,
         INCOMECD15
         )

# join requires the same case for character strings
data_grants$institution <- tolower(data_grants$institution)
data_museums$COMMONNAME <- tolower(data_museums$COMMONNAME)

# join by name, inner join retains only matching rows, and filter for last 10 years
data <- data_grants %>%
  inner_join(data_museums,
            by = c("institution" = "COMMONNAME"),
            relationship = "many-to-many"
            ) %>%
  filter(year >= 2013)
```

```{r include=FALSE}

# count, total funds, and mean award over last 10 years
data %>%
  summarize(grants = n(),
            total_funds = sum(funds),
            mean_award = mean(funds)
            )
```

```{r echo=FALSE}

# plot of total funds each year
data %>%
  group_by(year) %>%
  summarize(funds = sum(funds)) %>%
  ggplot(aes(x = year,
             y = funds
             )
         ) +
  geom_line() +
  scale_x_continuous("Year",
                     breaks = c(2013, 2014, 2015, 2016, 2017,
                                2018, 2019, 2020, 2021, 2022),
                     minor_breaks = NULL
                     ) +
  scale_y_continuous("Grant Funds",
                     labels = scales::label_dollar(),
                     limits = c(0, 15000000)
                     ) +
  theme_classic()
```

------------------------------------------------------------------------

<br>

### Where did all that funding go?

The 628 grants were awarded to 254 different museums. Of those 254, 44% received one grant and the remaining 56% received multiple grants. The American Museum of Natural History in New York tops the list both in terms of number of grants received (12) and the total funding received (\$3,411,368).

```{r include=FALSE}

data %>%
  count(institution, sort = TRUE) %>%
  summarize(institutions = n())
```

```{r include=FALSE}

data %>%
  count(institution) %>%
  group_by(n) %>%
  summarize(count = n()) %>%
  arrange(n)
```

```{r include=FALSE}

data %>%
  group_by(institution) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
#  arrange(-count)
  arrange(-funds)
```

```{r echo=FALSE}

# plot of number of grants
data %>%
  count(institution) %>%
  group_by(n) %>%
  summarize(count = n()) %>%
  arrange(n) %>%
  ggplot(aes(x = n,
             y = count
             )
         ) +
  geom_col() +
  scale_x_continuous("Number of Grants Received",
                     breaks = c(0:12),
                     minor_breaks = NULL
                     ) +
  scale_y_continuous("Number of Museums",
                     limits = c(0, 120)
                     ) +
  theme_classic()
```

<br>

At the state level, museums in New York, Illinois, and California received the most funding. There weren't any museums in South Dakota or Oklahoma that were awarded grants from the IMLS.

```{r eval=FALSE, include=FALSE}

# number of grants
data %>%
  group_by(state) %>%
  summarize(grants = n(),
            total_funds = sum(funds),
            mean_award = mean(funds)
            ) %>%
  arrange(-total_funds)
```

```{r eval=FALSE, include=FALSE}

# number of museums in each state
data_museums_raw %>%
  group_by(ADSTATE) %>%
  summarize(count = n()) %>%
  arrange(-count)
```

```{r message=FALSE, warning=FALSE, include=FALSE}

# import state abbreviations
state_abbr <- read_csv("state_abbr.csv")

# change state name to lower case
state_abbr$state <- tolower(state_abbr$state)

# summarize data to state and total funds
grants_state <- data %>%
  group_by(state) %>%
  summarize(total_funds = sum(funds)) %>%
  rename(state_abbr = state) %>%
  left_join(state_abbr,
            by = c("state_abbr" = "abbreviation")
            )

# for map data (from R Graphics Cookbook)
states_map <- map_data("state")

# join
grants_map <- states_map %>%
  left_join(grants_state,
            by = c("region" = "state")
            )
```

```{r echo=FALSE}

# plot map
grants_map %>%
  ggplot(aes(x = long,
             y = lat,
             group = group,
             fill = total_funds
             )
         ) +
  geom_polygon(color = "black") +
  coord_map("mercator") +
  scale_fill_viridis_c(labels = scales::label_dollar()) +
  labs(fill = "Grant Funding") +
  theme_void()
```

```{r eval=FALSE, include=FALSE}

grants_map %>%
  ggplot(aes(map_id = region,
             fill = total_funds
             )
         ) +
  geom_map(map = states_map,
           color = "black"
           ) +
  expand_limits(x = states_map$long,
                y = states_map$lat
                ) +
  coord_map("mercator") +
  # scale_fill_gradient2(low = "#559999",
  #                      mid = "grey90",
  #                      high = "#BB650B",
  #                      midpoint = 915738
  #                      ) +
  scale_fill_viridis_c() +
  labs(fill = "Grant Funding") +
  theme_void()
```

<br>

There's a clear leader when comparing funding by museum discipline. Nearly \$30 million was awarded to art museums, almost 50% more than the runner up, children's museums. Botanical gardens, zoos/aquariums, and science museums received similar amounts. History museums received the least amount of funds with \$6,201,416.

```{r eval=FALSE, include=FALSE}

# absolute funds
data %>%
  group_by(DISCIPL) %>%
  summarize(funds = sum(funds)) %>%
  arrange(-funds)
```

```{r echo=FALSE}

# summarize to funds by discipline, save as new object
data_disciplines <- data %>%
  group_by(DISCIPL) %>%
  summarize(funds = sum(funds))

# rename discipline factor levels
# levels(data_disciplines$DISCIPL) <- list(ART = "Art",
#                                          BOT = "Botanical Garden",
#                                          CMU = "Children's",
#                                          HST = "History",
#                                          NAT = "Natural History",
#                                          SCI = "Science",
#                                          ZAW = "Zoo/Aquarium"
#                                          )

# rename discipline factor levels
levels(data_disciplines$DISCIPL) <- c("Art",
                                      "Botanical",
                                      "Children's",
                                      "History",
                                      "Natural History",
                                      "Science",
                                      "Zoo/Aquarium"
                                      )
```

```{r echo=FALSE}

# plot of absolute funds by discipline
data_disciplines %>%
  ggplot(aes(x = reorder(DISCIPL, -funds),
             y = funds,
             # fill = DISCIPL
             )
         ) +
  geom_col() +
  scale_x_discrete("Museum Discipline") + # change x-axis label
  scale_y_continuous("Grant Funds", # change y-axis label
                   labels = scales::label_dollar() # reformat y-axis to $ with commas
                   ) +
  guides(fill = "none") + # remove legend
  theme_classic()
  # theme(axis.text.x = element_text(angle = 45)) # rotate x-axis labels
```

```{r eval=FALSE, include=FALSE}

data_disciplines %>%
  ggplot(aes(x = reorder(DISCIPL, -funds),
             y = funds,
             fill = DISCIPL
             )
         ) +
  geom_col() +
  scale_y_continuous("Grant Funds", # change y-axis label
                   labels = scales::label_dollar() # reformat y-axis to $ with commas
                   ) +
  labs(x = "Museum Discipline") + # change x-axis label
  geom_text(aes(label = DISCIPL,
                vjust = -0.8,
                hjust = 0.3
                ),
            angle = 45
            ) +
  theme_classic()
```

------------------------------------------------------------------------

<br>

### More funding went to museums with higher income

I was curious to see how the grants are distributed relative to income reported by the museums. Surprisingly, more funding went to museums with higher income levels. This contradicted my assumption that grants would go to institutions that needed funding, and that museums with higher income need less funding because they have the resources to support their initiatives.

```{r echo=FALSE}

# plot of absolute funds
data %>%
  filter(!is.na(INCOMECD15)) %>%
  group_by(INCOMECD15) %>%
  summarize(funds = sum(funds)) %>%
  add_row(INCOMECD15 = as.factor(1),
          funds = 0
          ) %>%
  ggplot(aes(x = INCOMECD15,
             y = funds,
             fill = INCOMECD15
             )
         ) +
  geom_col() +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Grant Funds", # change y-axis label
                   labels = scales::label_dollar() # reformat y-axis to $ with commas
                   ) +
  scale_fill_discrete("Income Level", 
                      labels =  c("$0", 
                               "$1 - $9,999", 
                               "$10,000 - $24,999", 
                               "$25,000 - $99,999",
                               "$100,000 - $499,999", 
                               "$500,000 - $999,999",
                               "$1,000,000 - $4,999,999", 
                               "$5,000,000 - $9,999,999",
                               "$10,000,000 - $49,999,999", 
                               "$50,000,000 to greater"
                               )
                    ) +
  theme_classic()
```

*Note - income data wasn't available for 15% of the museums and they aren't included in the chart above.*

------------------------------------------------------------------------

<br>

### But how does this compare to income levels across *all* museums?

Before diving deeper it's worth comparing income levels of museums that received a grant to income levels of all museums in the IMLS data. By including the **6,726 museums that did not receive a grant** over the last 10 years [FOOTNOTE]{.underline} the relative proportions for each income level are clear. There are **nearly 1,300 museums that reported zero income**, the largest group in the data. There's a "bulge" in the middle income levels and another spike at the highest level, with **689 museums reporting over \$50 million in income**. With this context it still appears that more grants were awarded to museums with higher income levels.

```{r include=FALSE}

# create object for museums that did not receive a grant
data_not_awarded <- data_museums %>%
  full_join(data_grants, 
            by = c("COMMONNAME" = "institution"), 
            keep = TRUE,
            relationship = "many-to-many"
            ) %>%
  filter(!is.na(COMMONNAME)) %>%    # remove awarded institutions not in the museum list
  replace_na(list(year = 9999,      # replace NAs with values for analysis
                  funds = 0
                  )
             ) %>%
  mutate(awarded = if_else(funds > 0, TRUE, FALSE)) %>%
  filter(year >= 2013) %>%
  filter(awarded == FALSE)

# count not awarded
data_not_awarded_count <- data_not_awarded %>%
  filter(!is.na(INCOMECD15)) %>%
  group_by(INCOMECD15) %>%
  summarize(count = n()) %>%
  mutate(awarded = FALSE)

# count awarded
data_awarded_count <- data %>%
  filter(!is.na(INCOMECD15)) %>%
  group_by(institution) %>%
  slice_head() %>%
  ungroup() %>%
  group_by(INCOMECD15) %>%
  summarize(count = n()) %>%
  add_row(INCOMECD15 = as.factor(1),
          count = 0
          ) %>%
  mutate(awarded = TRUE) %>%
  arrange(INCOMECD15)

# combine
data_count <- data_awarded_count %>%
  bind_rows(data_not_awarded_count)
```

```{r eval=FALSE, include=FALSE}

data_count %>%
  group_by(INCOMECD15) %>%
  summarize(number = sum(count))
```

```{r echo=FALSE}

data_count %>%
  ggplot(aes(x = INCOMECD15,
             y = count,
             fill = awarded
             )
         ) +
  geom_col(position = "stack") +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Number of Museums") +
  scale_fill_manual("Received grant?", 
                      labels =  c("No",
                                  "Yes"
                                  ),
                      values = c("gray",
                                 "red")
                      ) +
  theme_classic()
```

*Note - income data wasn't available for 35% of the museums that did not receive a grant and they aren't included in the chart above.*

------------------------------------------------------------------------

<br>

### Most grant programs require matching funds

Let's look at how these grants are actually awarded. The IMLS has multiple grant programs with various eligibility criteria, funding amounts, and cost share requirements. In this data, the top four programs by funding awarded **represent 95% of the total funding** over the last 10 years. Of those, the top three require that the museum, "provide funds from non-federal sources in an amount that is equal to or greater than" the amount requested in the grant application. For example, if a museum applies for a \$100,000 grant through the *Museums for America* program they must match that amount through other (non-federal) means.

<br>

```{r echo=FALSE}

program_table <- tibble("Rank" = c(1:4),
                        "Program" = c("Museums for America",
                                      "National Leadership Grants - Museums",
                                      "Museums Empowered",
                                      "Inspire! Grants for Small Museums"),
                        "Total Funds" = c("$70,578,719",
                                          "$16,292,260",
                                          "$7,966,050",
                                          "$2,141,872"
                                          ),
                        "Number of Grants" = c(465,
                                               37,
                                               53,
                                               46
                                               ),
                        "Requires Matching Funds?" = c("Yes",
                                                       "Yes",
                                                       "Yes",
                                                       "No")
                           )

program_table %>%
  kbl() %>%
  kable_classic(full_width = FALSE,
                html_font = "helvetica"
                )
```

<br>

Isolating each program confirms that the top three programs skew towards higher income levels. The *Inspire! Grants for Small Museums* program, which does not require matching funds, was awarded to museums with mid-range and lower income levels. FOOTNOTE

```{r echo=FALSE}

p1 <- data %>%
  filter(program == "Museums for America") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(1),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = funds
             )
         ) +
  geom_col() +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Grant Funds", # change y-axis label
                     labels = scales::label_dollar(scale_cut = scales::cut_short_scale()),
                     limits = c(0, 22000000)
                     ) +
  labs(title = "Museums for America") +
  theme_classic() +
  theme(plot.title = element_text(size = rel(0.9),
                                  face = "bold"
                                  ),
        axis.title = element_text(size = rel(0.9))
        )

p2 <- data %>%
  filter(program == "National Leadership Grants - Museums") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(0, 1, 2, 3, 3)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = funds
             )
         ) +
  geom_col() +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Grant Funds", # change y-axis label
                     labels = scales::label_dollar(scale_cut = scales::cut_short_scale()),
                     limits = c(0, 22000000)
                     ) +
  labs(title = "Nat'l Leadership Grants - Museums") +
  theme_classic() +
  theme(plot.title = element_text(size = rel(0.9),
                                  face = "bold"
                                  ),
        axis.title = element_text(size = rel(0.9))
        )

p3 <- data %>%
  filter(program == "Museums Empowered: Professional Development Opportunities for Museum Staff") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(1, 2, 3, 3)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = funds
             )
         ) +
  geom_col() +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Grant Funds", # change y-axis label
                     labels = scales::label_dollar(scale_cut = scales::cut_short_scale()),
                     limits = c(0, 22000000)
                     ) +
  labs(title = "Museums Empowered") +
  theme_classic() +
  theme(plot.title = element_text(size = rel(0.9),
                                  face = "bold"
                                  ),
        axis.title = element_text(size = rel(0.9))
        )

p4 <- data %>%
  filter(program == "Inspire! Grants for Small Museums") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(1, 2, 7, 8, 9)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = funds
             )
         ) +
  geom_col() +
  scale_x_discrete("Income Level") +
  scale_y_continuous("Grant Funds", # change y-axis label
                     labels = scales::label_dollar(scale_cut = scales::cut_short_scale()),
                     limits = c(0, 22000000)
                     ) +
  labs(title = "Inspire! Grants for Small Museums") +
  theme_classic() +
  theme(plot.title = element_text(size = rel(0.9),
                                  face = "bold"
                                  ),
        axis.title = element_text(size = rel(0.9))
        )

(p1 | p2) / (p3 | p4)
```

------------------------------------------------------------------------

<br>

### The burden of matching funds

Placeholder.

```{r}

# test with Museums for America only to start
# median grant value through Museums for America
data %>%
  filter(program == "Museums for America") %>%
  summarize(median_grant = median(funds)) 
```

```{r}

# create plot of income codes
income_codes <- tibble(
  code = as.factor(c(0:9)),
  low = c(0, 1, 10000, 25000, 100000, 500000, 1000000, 5000000, 10000000, 50000000),
  high = c(0, 9999, 24999, 99999, 499999, 999999, 4999999, 9999999, 49999999, 50000000)
) %>%
  mutate(mid = ((high - low)/2) + low,
         median_grant = 149967,
         ratio = mid/median_grant,
         percent = (median_grant/mid) * 100
         )

income_codes
```

```{r}

income_codes %>%
  ggplot(aes(x = code,
             y = mid
             )
         ) +
  geom_col() +
  geom_point(aes(x = code,
                 y = median_grant
                 )
             )
```

```{r}

# ration of income (mid) to median grant
income_codes %>%
  ggplot(aes(x = code,
             y = ratio
             )
         ) +
  geom_col()
```

------------------------------------------------------------------------

<br>

### Call to action

-   Access to museums improve outcomes

-   IMLS should direct more funding to areas that need it

-   People can donate to the least likely museums

------------------------------------------------------------------------

<br>

### Footnotes

1.  I don't know why the institutions did not receive a grant.
2.  INCOMECD15 used as benchmark for income even though it could be different before or after.
3.  Institutions may have received a grant in a different year than when their INCOMECD15 year
4.  The [application information](https://imls.gov/grants/available/inspire-grants-small-museums) for this program explains that it is designed to support small museums, but it's not clear how "small" is defined. If there is a correlation between size and income level, then museums with high income levels may not be eligible for this program. That could be one explanation for the distribution above. Unfortunately, the IMLS data doesn't contain information regarding size.

<br>

### Methodology

Placeholder.

The IMLS maintains a list of museums in the US in a set of files called the Museum Data Files (MDF). The MDF contain basic institutional identifying information for about 30,000 museums and related organizations in the US. They are split into three files using the following categories. The MDF was last updated in 2018.

The IMLS also makes available data regarding the grants they award to these institutions for various initiatives.

<br>

### Sources

Placeholder.

<br>

### To Do

-   CSS to set width of html output?
-   Link to GitHub repository
-   AK and HI in map
-   Color scale in map
-   Account for number of museums in state, discipline?
-   colors for funds by discipline
-   rename income Level or use a legend early on?
-   link to footnotes
-   bold key text
-   center charts
-   smart brevity?

```{r eval=FALSE, include=FALSE}

# SANDBOX

# institutions with more revenue/income can pay for staff who apply for grants

# impact of private/other funding on applying for grants, winning

```

```{r eval=FALSE, include=FALSE}

p1 <- data %>%
  filter(program == "Museums for America") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(1),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = count
             )
         ) +
  geom_col() +
  ylim(0, 150) +
  theme_classic()

p2 <- data %>%
  filter(program == "National Leadership Grants - Museums") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(0, 1, 2, 3, 3)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = count
             )
         ) +
  geom_col() +
  ylim(0, 150) +
  theme_classic()

p3 <- data %>%
  filter(program == "Museums Empowered: Professional Development Opportunities for Museum Staff") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(1, 2, 3, 3)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = count
             )
         ) +
  geom_col() +
  ylim(0, 150) +
  theme_classic()

p4 <- data %>%
  filter(program == "Inspire! Grants for Small Museums") %>%
  group_by(INCOMECD15) %>%
  filter(!is.na(INCOMECD15)) %>%
  summarize(count = n(),
            funds = sum(funds)
            ) %>%
  add_row(INCOMECD15 = as.factor(c(1, 2, 7, 8, 9)),
          count = 0,
          funds = 0
          ) %>%
  arrange(INCOMECD15) %>%
  ggplot(aes(x = INCOMECD15,
             y = count
             )
         ) +
  geom_col() +
  ylim(0, 150) +
  theme_classic()

(p1 | p2) / (p3 | p4)
```

```{r}

data %>%
  group_by(program) %>%
  summarize(total_funds = sum(funds)) %>%
  arrange(-total_funds)
```
